 

### 2020-09-28 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

**The strength of this buffering varies across forests, being influenced by both forest characteristics and the biophysical environment.**
First, buffering increases with canopy cover.
* greater cover --> lower max T and VPD, higher minT  [@davis_microclimatic_2019]
* greater cover -->  lower max T and light availability [@zellweger_seasonal_2019]
* [@thom_effects_2020; Tymen et al. 2017 ; Jucker et al. 2018 ; Hardwick et al. 2015]
*Presumably, buffering would also be affected by canopy roughness, which affects turbulent air flow and the canopy boundary layer.*
Taller trees don’t necessarily increase buffering [@zellweger_seasonal_2019]. 
Species that have higher shade casting ability increase T buffering (Zelllweger et al. 2019) [@zellweger_seasonal_2019]
The strength of buffering also varies with respect to geographic and climatic factors.
* Distance to coast, topographic position, elevation [@zellweger_seasonal_2019]
* [@davis_microclimatic_2019]

@@

```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below (easiest in alphabetical order)
```
---
nocite: | 
  @mau_temperate_2018
---

@@

>>>>

**The strength of this buffering varies across forests, being influenced by both forest characteristics and the biophysical environment.**
First, buffering increases with canopy cover.
* greater cover --> lower max T and VPD, higher minT  [@davis_microclimatic_2019]
* greater cover -->  lower max T and light availability [@zellweger_seasonal_2019]
* [@thom_effects_2020; Tymen et al. 2017 ; Jucker et al. 2018 ; Hardwick et al. 2015]
*Presumably, buffering would also be affected by canopy roughness, which affects turbulent air flow and the canopy boundary layer.*
Taller trees don’t necessarily increase buffering [@zellweger_seasonal_2019]. 
Species that have higher shade casting ability increase T buffering (Zelllweger et al. 2019) [@zellweger_seasonal_2019]
The strength of buffering also varies with respect to geographic and climatic factors.
* Distance to coast, topographic position, elevation [@zellweger_seasonal_2019]
* [@davis_microclimatic_2019]
However vertical climate gradients within forests are known to be steeper than those driven by elevation and latitude [@nakamuraForestsTheirCanopies2017,  @martinCovarianceSunShade2020a]

@@

```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below (easiest in alphabetical order)
```
---
nocite: | 
  @mau_temperate_2018, @cobleHowVerticalPatterns2016a, @slotPhotosyntheticHeatTolerance2019, @valladaresShadeToleranceKey2008b, @beaumontVerticalGradientsLeaf2009, @sackHowStrongIntracanopy2006,@cobleLightDrivesVertical2014, @scartazzaInvestigatingEuropeanBeech2016, @martinCovarianceSunShade2020, @bolstad_foliar_1999, @harleyEffectsLightTemperature1996, @kullDistributionLeafPhotosynthetic1998, @kafutiFoliarWoodTraits2020, @duursmaVerticalCanopyGradients2006, @vanwittenbergheVariabilityStomatalConductance2012, 
---

@@

[@nakamuraForestsTheirCanopies2017]--> modelling ideas?
